---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Reading in Mr. Trash Wheel excel file 
  Importing file
    * specifying range
    * excluding rows with images
  cleaning names
  filter out rows with monthly totals
    * just want dumpster information
  Round sports_balls to an integer




 mutate(sports_balls = as.integer(trash_df$sports_balls))

## Read and clean precipitation data for 2018 & 2019

Importing 2018 and 2019 precipitation sheets

  Cleaning and renaming data 

```{r}
precip_2018_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 5, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018") %>% 
  select(year, everything())
  

precip_2019_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 4,range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
     year = "2019") %>% 
  select(year, everything())
```



## Joining 2019 and 2018 data
  Using left join
  
```{r}
precip_df = 
  left_join(precip_2018_df, precip_2019_df, by = "month")
```




    



