---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

## Reading in Mr. Trash Wheel excel file 
  Importing file
* specifying range
* excluding rows with images
* cleaning names
* filter out rows with monthly totals
** just want dumpster information
* Round sports_balls to an integer

```{r}
trash_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 1, range = "A2:N408") %>%
 janitor::clean_names() %>% 
  drop_na() %>% 
   mutate(
      sports_balls = as.integer(sports_balls))
```


## Read and clean precipitation data for 2018 & 2019

Importing 2018 and 2019 precipitation sheets

  Cleaning and renaming data 

```{r}
precip_2018_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 5, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018") %>% 
   mutate(
     month = as.integer(month)) %>% 
  select(year, everything())
  

precip_2019_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 4,range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
     year = "2019") %>% 
   mutate(
     month = as.integer(month)) %>% 
  select(year, everything())
```


## Joining 2019 and 2018 data
  Using left join
  
```{r}
precip_df = 
  bind_rows(precip_2018_df, precip_2019_df) %>% 
  mutate(
      month = month.name[month])
  
```

There are `r nrow(trash_df)` observations in `trash_df` and the variables are `r names(trash_df)`. There are `nrow(precip_df)` observations in `precip_df`, which combines the precipitation data from 2018 and 2019. The varaibles in this combined data set are `r names(precip_df)`. 

The median number of sports balls in a dumpster in 2019 was `r median(pull(trash_df, sports_balls))`. The total precipitation in 2018 was `r sum(pull(precip_2018_df, total))`. 

### insert paragraph using in line code
    
# Problem 2 

Importing and cleaning pols-month, unemployment, and snp

 Starting with pols-month
 

```{r}
pols_df = read.csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year","month", "day")) %>%  
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.abb[month]) %>% 
  unite('president', prez_dem,prez_gop, sep = "") %>% 
  mutate(
      president = as.numeric(president)) %>% 
  mutate(president = recode(president, '10' = 'dem', '01' = 'gop')) %>% 
   mutate(month = tolower(month)) %>% 
  select(-day)
```

 Cleaning snp
1.  reformatting date
2.  separating date into day, month, year (removing day)
3. recoding month into month names


```{r}
library(lubridate)

snp_df = read.csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.abb[month]) %>% 
   mutate(month = tolower(month)) %>% 
  select(-day)
```


Tidy unemployment
* cleaning
* switching from wide to long format

```{r}
unemployment_df = read.csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemp_perc") %>% 
    mutate(year = as.character(year))
```

Merging data sets
1. starting with snp & pols

```{r}
pols_snp_df = 
    right_join(pols_df, snp_df, by = c("year", "month"))
```

Merging unemployment into pols_snp

## Joining all three data sets

```{r}
combined_df =
  left_join(pols_snp_df, unemployment_df, by = c("year", "month")) 
```

There are `r nrow(combined_df)` observations in `combined_df` and the variables are `r names(combined_df)`. The range of `years` is `r range(pull(combined_df, year))`.



# Problem 3

Loading popular baby names data set
1. Cleaning data
2. Formating all values in lower case
3. removing duplicates

```{r}
baby_df = read.csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  mutate(child_s_first_name = tolower(child_s_first_name),
         gender = tolower(gender),
         ethnicity = tolower(ethnicity)) %>%
  distinct(child_s_first_name, year_of_birth, .keep_all = TRUE)
```

Making a table
1. Ranking popularity of the name "Olivia"
  ** rows for ethnicity
  ** columns for year

Reorder years
Error with ethnicity name
```{r}
baby_olivia_df = 
   filter(baby_df, child_s_first_name == "olivia", gender == "female") %>% 
  select(-child_s_first_name, -gender, -count) %>%
 pivot_wider(
   names_from = "year_of_birth",
   values_from = "rank") %>% 
  relocate("ethnicity", "2011", "2012", "2013", "2014", "2015", "2016")
```
2. Making table for popular "boy" names over time 

```{r}
popular_names_male_df = 
  subset(baby_df, gender == "male") %>% 
  select(child_s_first_name, year_of_birth, rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% 
   relocate("child_s_first_name", "2011", "2012", "2013", "2014", "2015", "2016")
```


Making a scatter plot for white non-hispanic children born in 2016
1. Y- axis = number of children with a name
2. X-axis = rank in popularity of that name
  
```{r}
white_boy_names_df =
  filter(baby_df, gender != "female", ethnicity == "white non hispanic", year_of_birth == 2016)

ggplot(white_boy_names_df, aes(x = rank, y = count)) +
  geom_point()

```

